---
title: "Week 2 - Homework"
author: "STAT 420, Summer 2020, Hyunjoon Rhee"
date: ''
output:
  html_document: 
    theme: readable
    toc: yes  
  pdf_document: default
urlcolor: cyan
---


***

## Exercise 1 (Using `lm`)

For this exercise we will use the `cats` dataset from the `MASS` package. You should use `?cats` to learn about the background of this dataset.

**(a)** Suppose we would like to understand the size of a cat's heart based on the body weight of a cat. Fit a simple linear model in `R` that accomplishes this task. Store the results in a variable called `cat_model`. Output the result of calling `summary()` on `cat_model`.
```{r}
cats = MASS::cats
?cats
#View(cats)
cat_model = lm(Hwt ~ Bwt, data = cats)
summary(cat_model)
```


**(b)** Output only the estimated regression coefficients. Interpret $\hat{\beta_0}$ and $\beta_1$ in the *context of the problem*. Be aware that only one of those is an estimate.
```{r}
coef(cat_model)
```
$\hat{\beta_0}$ is the estimated value of intercept. Whereas $\beta_1$ is just the slope of the data.

**(c)** Use your model to predict the heart weight of a cat that weights **3.1** kg. Do you feel confident in this prediction? Briefly explain.
```{r}
predict(cat_model, newdata = data.frame(Bwt = 2.7))
```
Because the value is in between the range, it could be said that it is good to feel confident in the prediction.

**(d)** Use your model to predict the heart weight of a cat that weights **1.5** kg. Do you feel confident in this prediction? Briefly explain.
```{r}
predict(cat_model, newdata = data.frame(Bwt = 4.4))
```
Because the value is somewhat far from the main cluster, it could be said that it is not good to feel confident in the prediction.

**(e)** Create a scatterplot of the data and add the fitted regression line. Make sure your plot is well labeled and is somewhat visually appealing.
```{r}
plot(Hwt ~ Bwt, data = cats,
     xlab = "Cat Weight",
     ylab = "Heart Weight",
     main = "Cats body vs heart weight",
     pch  = 20,
     cex  = 2,
     col  = "darkorange"
     );
abline(cat_model, lwd = 3, col = "green");
```

**(f)** Report the value of $R^2$ for the model. Do so directly. Do not simply copy and paste the value from the full output in the console after running `summary()` in part **(a)**.
```{r}
x = cats$Bwt
y = cats$Hwt
Sxy = sum((x-mean(x)) * (y-mean(y)))
Sxx = sum((x-mean(x))^2)
Syy = sum((y-mean(y))^2)
beta_1_hat = Sxy/Sxx
beta_0_hat = mean(y) - beta_1_hat * mean(x)
y_hat = beta_0_hat + beta_1_hat * x
SST   = sum((y - mean(y)) ^ 2)
SSReg = sum((y_hat - mean(y)) ^ 2)
R2 = SSReg/SST
print(R2)
```


***

## Exercise 2 (Writing Functions)

This exercise is a continuation of Exercise 1.

**(a)** Write a function called `get_sd_est` that calculates an estimate of $\sigma$ in one of two ways depending on input to the function. The function should take three arguments as input:

- `fitted_vals` - A vector of fitted values from a model
- `actual_vals` - A vector of the true values of the response
- `mle` - A logical (`TRUE` / `FALSE`) variable which defaults to `FALSE`

The function should return a single value:

- $s_e$ if `mle` is set to `FALSE`.
- $\hat{\sigma}$ if `mle` is set to `TRUE`.

```{r}
get_sd_est = function(fitted_vals, actual_vals, mle = FALSE) {
  #fitted value = y hat
  e = actual_vals - fitted_vals
  n = length(e)
  s2_e = sum(e^2)/(n-2)
  if(mle == FALSE) {
    return(sqrt(s2_e))
  }
  sig_hat = sqrt(sum(e^2)/n)
  if(mle == TRUE) {
    return(sig_hat)
  }
}
```


**(b)** Run the function `get_sd_est` on the residuals from the model in Exercise 1, with `mle` set to `FALSE`. Explain the resulting estimate in the context of the model.
```{r}
get_sd_est(y_hat,y,FALSE)
```
This means that the difference between the estimated value of y and the mean y is usually about 1.42373.

**(c)** Run the function `get_sd_est` on the residuals from the model in Exercise 1, with `mle` set to `TRUE`. Explain the resulting estimate in the context of the model. Note that we are trying to estimate the same parameter as in part **(b)**.
```{r}
get_sd_est(y_hat,y,TRUE)
```
The difference between the estimated y and mean y is 1.442252. This has a biased estimate.

**(d)** To check your work, output `summary(cat_model)$sigma`. It should match at least one of **(b)** or **(c)**.
```{r}
summary(cat_model)$sigma
```
It matches with (b).

***

## Exercise 3 (Simulating SLR)

Consider the model

\[
Y_i = 5 + -3 x_i + \epsilon_i
\]

with 

\[
\epsilon_i \sim N(\mu = 0, \sigma^2 = 10.24)
\]

where $\beta_0 = 5$ and $\beta_1 = -3$.

This exercise relies heavily on generating random observations. To make this reproducible we will set a seed for the randomization. Alter the following code to make `birthday` store your birthday in the format: `yyyymmdd`. For example, [William Gosset](https://en.wikipedia.org/wiki/William_Sealy_Gosset), better known as *Student*, was born on June 13, 1876, so he would use:

```{r}
birthday = 19960926
set.seed(birthday)
```

**(a)** Use `R` to simulate `n = 25` observations from the above model. For the remainder of this exercise, use the following "known" values of $x$.

```{r}
x = runif(n = 25, 0, 10)
```
```{r}
sim_slr = function(x, beta_0 = 10, beta_1 = 5, sigma = 1) {
  n = length(x)
  epsilon = rnorm(n, mean = 0, sd = sigma)
  y = beta_0 + beta_1 * x + epsilon
  data.frame(predictor = x, response = y)
}
model = sim_slr(x, 5, -3, sqrt(10.24))
```

You may use [the `sim_slr ` function provided in the text](http://daviddalpiaz.github.io/appliedstats/simple-linear-regression.html#simulating-slr). Store the data frame this function returns in a variable of your choice. Note that this function calls $y$ `response` and $x$ `predictor`.

**(b)** Fit a model to your simulated data. Report the estimated coefficients. Are they close to what you would expect? Briefly explain.
```{r}
linearmodel = lm(response ~ predictor, data = model)
linearmodel
```
The estimated coefficients are very similar to the value of $\beta_0$ and $\beta_1$ which are 5 and -3 in the equation above. The values from the linear model, 5.186, and -3.263 are not totally off, but rather close to the value.

**(c)** Plot the data you simulated in part **(a)**. Add the regression line from part **(b)** as well as the line for the true model. Hint: Keep all plotting commands in the same chunk.
```{r}
library(ggplot2)
(ggplot(model, aes(x = predictor, y = response)) + geom_point(color = "green")
  + xlab("predictor")
  + ylab("response")
  + ggtitle("SLR simulate")
  + geom_abline(intercept = coef(linearmodel)["(Intercept)"], slope = coef(linearmodel)["predictor"]))
```


**(d)** Use `R` to repeat the process of simulating `n = 25` observations from the above model $1500$ times. Each time fit a SLR model to the data and store the value of $\hat{\beta_1}$ in a variable called `beta_hat_1`. Some hints:

- Consider a `for` loop.
- Create `beta_hat_1` before writing the `for` loop. Make it a vector of length $1500$ where each element is `0`.
- Inside the body of the `for` loop, simulate new $y$ data each time. Use a variable to temporarily store this data together with the known $x$ data as a data frame.
- After simulating the data, use `lm()` to fit a regression. Use a variable to temporarily store this output.
- Use the `coef()` function and `[]` to extract the correct estimated coefficient.
- Use `beta_hat_1[i]` to store in elements of `beta_hat_1`.
- See the notes on [Distribution of a Sample Mean](http://daviddalpiaz.github.io/appliedstats/introduction-to-r.html#distribution-of-a-sample-mean) for some inspiration.

You can do this differently if you like. Use of these hints is not required.
```{r}
beta_hat_1 = c(rep(0, 1500))
for(i in 1:1500) {
  x = runif(n = 25, 0, 10);
  model = sim_slr(x, 5, -3, sqrt(10.24));
  linearmodel = lm(response ~ predictor, data = model)
  beta_hat_1[i] = coef(linearmodel)["predictor"]
}
```

**(e)** Report the mean and standard deviation of `beta_hat_1`. Do either of these look familiar?
```{r}
mean(beta_hat_1,na.rm=TRUE)
sd(beta_hat_1,na.rm=TRUE)
```
The value of mean is very close to the original value of -3.

**(f)** Plot a histogram of `beta_hat_1`. Comment on the shape of this histogram.
```{r}
hist(beta_hat_1,
     main   = "", 
     xlab   = expression(hat(beta)[1]), 
     col    = "darkorange", 
     border = "dodgerblue")
```
The shape of the graph seems to be much like normal distribution with a mean of -3.
***

## Exercise 4 (Be a Skeptic)

Consider the model

\[
Y_i = 3 + 0 \cdot x_i + \epsilon_i
\]

with

\[
\epsilon_i \sim N(\mu = 0, \sigma^2 = 4)
\]

where $\beta_0 = 3$ and $\beta_1 = 0$.

Before answering the following parts, set a seed value equal to **your** birthday, as was done in the previous exercise.

```{r}
birthday = 19960926
set.seed(birthday)
```

**(a)** Use `R` to repeat the process of simulating `n = 75` observations from the above model $2500$ times. For the remainder of this exercise, use the following "known" values of $x$.

```{r}
x = runif(n = 75, 0, 10)
```

Each time fit a SLR model to the data and store the value of $\hat{\beta_1}$ in a variable called `beta_hat_1`. You may use [the `sim_slr ` function provided in the text](http://daviddalpiaz.github.io/appliedstats/simple-linear-regression.html#simulating-slr). Hint: Yes $\beta_1 = 0$.
```{r}
beta_hat_1 = rep(0, 2500)
for(i in seq_along(beta_hat_1)) {
  model2 = sim_slr(x, 3, 0, 2)
  linearmodel2 = lm(response ~ predictor, data = model2)
  beta_hat_1[i] = coef(linearmodel2)[2]
}
```

**(b)** Plot a histogram of `beta_hat_1`. Comment on the shape of this histogram.
```{r}
hist(beta_hat_1,
     main   = "", 
     xlab   = expression(hat(beta)[1]), 
     col    = "darkorange", 
     border = "dodgerblue")
```
The shape of the graph is almost like normal distribution graph, and has a mean of 0.

**(c)** Import the data in [`skeptic.csv`](skeptic.csv) and fit a SLR model. The variable names in `skeptic.csv` follow the same convention as those returned by `sim_slr()`. Extract the fitted coefficient for $\beta_1$.
```{r}
skeptic = read.csv("skeptic.csv")
model3 = lm(response ~ predictor, data = skeptic)
beta_1 = coef(model3)["(Intercept)"]
beta_0 = coef(model3)["predictor"]
beta_1
beta_0
```

**(d)** Re-plot the histogram from **(b)**. Now add a vertical red line at the value of $\hat{\beta_1}$ in part **(c)**. To do so, you'll need to use `abline(v = c, col = "red")` where `c` is your value.
```{r}
hist(beta_hat_1,
     main   = "", 
     xlab   = expression(hat(beta)[1]), 
     col    = "darkorange", 
     border = "dodgerblue")
abline(v = coef(model3)["predictor"], col = "blue")
```


**(e)** Your value of $\hat{\beta_1}$ in **(c)** should be negative. What proportion of the `beta_hat_1` values is smaller than your $\hat{\beta_1}$? Return this proportion, as well as this proportion multiplied by `2`.
```{r}
mean(beta_hat_1 > coef(model3)[2])
2 * mean(beta_hat_1 > coef(model3)[2])
```


**(f)** Based on your histogram and part **(e)**, do you think the [`skeptic.csv`](skeptic.csv) data could have been generated by the model given above? Briefly explain.

The blue line above appears in the far left corner of the graph. This means that it is possible that the data could have generated the model, but it is highly unlikely to do so.

***

## Exercise 5 (Comparing Models)

For this exercise we will use the `Ozone` dataset from the `mlbench` package. You should use `?Ozone` to learn about the background of this dataset. You may need to install the `mlbench` package. If you do so, do not include code to install the package in your `R` Markdown document.

For simplicity, we will perform some data cleaning before proceeding.

```{r}
data(Ozone, package = "mlbench")
Ozone = Ozone[, c(4, 6, 7, 8)]
colnames(Ozone) = c("ozone", "wind", "humidity", "temp")
Ozone = Ozone[complete.cases(Ozone), ]
```

We have:

- Loaded the data from the package
- Subset the data to relevant variables
    - This is not really necessary (or perhaps a good idea) but it makes the next step easier
- Given variables useful names
- Removed any observation with missing values
    - This should be given much more thought in practice

For this exercise we will define the "Root Mean Square Error" of a model as

\[
\text{RMSE} = \sqrt{\frac{1}{n} \sum_{i = 1}^{n}(y_i - \hat{y}_i)^2}.
\]

**(a)** Fit three SLR models, each with "ozone" as the response. For the predictor, use "wind speed," "humidity percentage," and "temperature" respectively. For each, calculate $\text{RMSE}$ and $R^2$. Arrange the results in a markdown table, with a row for each model. Suggestion: Create a data frame that stores the results, then investigate the `kable()` function from the `knitr` package.
```{r}
windmodel = lm(ozone ~ wind, data = Ozone)
humiditymodel = lm(ozone ~ humidity, data = Ozone)
tempmodel = lm(ozone ~ temp, data = Ozone)

y = Ozone$ozone
n = length(y)

wind_y_hat = coef(windmodel)["(Intercept)"] + coef(windmodel)["wind"] * Ozone$wind
humidity_y_hat = coef(humiditymodel)["(Intercept)"] + coef(humiditymodel)["humidity"] * Ozone$humidity
temp_y_hat = coef(tempmodel)["(Intercept)"] + coef(tempmodel)["temp"] * Ozone$temp

SST = sum((y - mean(y))^2)
windSSReg = sum((wind_y_hat - mean(y))^2)
humidSSReg = sum((humidity_y_hat - mean(y))^2)
tempSSReg = sum((temp_y_hat - mean(y))^2)
windR2 = windSSReg / SST
humidR2 = humidSSReg / SST
tempR2 = tempSSReg / SST
windRSME = sqrt((1/n)*sum((y-wind_y_hat)^2))
humidRSME = sqrt((1/n)*sum((y-humidity_y_hat)^2))
tempRSME = sqrt((1/n)*sum((y-temp_y_hat)^2))

RMSElist = c("Wind" = windRSME, "Humidity" = humidRSME, "Temp" = tempRSME)
R2list = c("Wind" = windR2, "Humidity" = humidR2, "Temp" = tempR2)
result = data.frame(RMSElist, R2list)

library("knitr")
kable(result)
```

**(b)** Based on the results, which of the three predictors used is most helpful for predicting ozone readings? Briefly explain.

Temp predictor is the most helpful one to predict ozone reading. The R2 value is the highest and RMSE is the lowest. Low RMSE value means low errors of the model.

***
